

## Coursera: Intro to Computational Finance and Financial Econometrics: Assignment 2: Probability Calculations in R


1\. Suppose X is a normally distributed random variable with mean 0.05 and
variance (0.10)2. Compute the following: 

Pr(X > 0.10)


```{r}

    mu_x = 0.05
    sigma_x = 0.10
    
    # Pr(X > 0.10)
    1 - pnorm(0.10, mean=mu_x, sd=sigma_x)

```

2\. Pr(X < −.10)


```{r}
    
    pnorm(-0.10, mean=mu_x, sd=sigma_x)

```

3\. Pr(−0.05 < X < 0.15)

    
```{r}

    pnorm(0.15, mean=mu_x, sd=sigma_x) - pnorm(-0.05, mean=mu_x, sd=sigma_x)

```

4\. 1% quantile, q.01


```{r}

    qnorm(0.01, mean=mu_x, sd=sigma_x)

```

5\. 5% quantile, q.05


```{r}

    qnorm(0.05, mean=mu_x, sd=sigma_x)

```


6\. 95% quantile, q.95


```{r}

    qnorm(0.95, mean=mu_x, sd=sigma_x)

```

7\. 99% quantile, q.99


```{r}

    qnorm(0.99, mean=mu_x, sd=sigma_x)
    
```

8\. Let X denote the monthly return on Microsoft Stock and let Y denote the
monthly return on Starbucks stock. Assume that X∼N(0.05,(0.10)^2) and
Y∼N(0.025,(0.05)^2). 

Using a grid of values between –0.25 and 0.35, plot the normal curves for X and
Y. Make sure that both normal curves are on the same plot.

Which stock has the higher risk? ANSWER: MSFT, because it has a higher SD.


```{r}

    library(ggplot2)

    t = seq(-0.25, 0.35, 0.001)

    # MSFT stock
    mu_x = 0.05
    sigma_x = 0.10
    X = dnorm( t, mean=mu_x, sd=sigma_x )

    # SBUX stock
    mu_y = 0.025
    sigma_y = 0.05
    Y = dnorm( t, mean=mu_y, sd=sigma_y )

    ggplot() + 
        geom_line(aes(x=t, y=X), color="blue") + 
        geom_line(aes(x=t, y=Y), color="red")

```

9\. Let R denote the simple monthly return on Microsoft stock and let W0 denote
the initial wealth to be invested over the month. Assume that
R∼N(0.04,(0.09)2)) and that W0=$100,000

Determine the 1% and 5% value-at-risk (VaR) over the month on the
investment. That is, determine the loss in investment value that may occur over
the next month with 1% probability and with 5% probability.

```{r}

    mu_r = 0.04
    sigma_r = 0.09
    W0 = 1e5

    # 1% VAR = q_.01(R) * W0
    # 5% VAR = q_.05(R) * W0

    qnorm(0.01, mean=mu_r, sd=sigma_r) * W0
    qnorm(0.05, mean=mu_r, sd=sigma_r) * W0

```

10\. Let the r denote the continuously compounded monthly return on Microsoft
stock and let W0 denote the initial wealth to be invested over the month.
Assume that r∼iid N(0.04,(0.09)2) and that W0=$100,000.

Determine the 1% and 5% value-at-risk (VaR) over the month on the investment.
That is, determine the loss in investment value that may occur over the next
month with 1% probability and with 5% probability. (Hint: compute the 1% and 5%
quantile from the Normal distribution for r and then convert continuously
compounded return quantile to a simple return quantile using the transformation
R=er−1)


```{r}
    
    mu_r_cc = 0.04
    sigma_r_cc = 0.09

    # 1% VAR = (exp( q_.01(r_cc) ) - 1) * W0
    # 5% VAR = (exp( q_.05(r_cc) ) - 1 ) * W0

    W0 * ( exp( qnorm(0.01, mean=mu_r_cc, sd=sigma_r_cc) ) - 1 )
    W0 * ( exp( qnorm(0.05, mean=mu_r_cc, sd=sigma_r_cc) ) - 1 )


```

11\. Consider a one month investment in two Northwest stocks: Amazon and
Costco. Suppose you buy Amazon and Costco at the end of September at
PA,t−1=$38.23, PC,t−1=$41.11 and then sell at the end of October for
PA,t=$41.29, PC,t=$41.74. (Note: these are actual closing prices for 2004 taken
from Yahoo!)

What are the simple monthly returns for the two stocks?


```{r}

    pA0 = 38.23
    pA1 = 41.29

    pC0 = 41.11
    pC1 = 41.74

    # simple return: (pt1 - pt0)/pt0
    R_A = (pA1 - pA0) / pA0
    R_C = (pC1 - pC0) / pC0

    R_A
    R_C


```

12\. What are the continuously compounded returns for the two stocks?


```{r}

    # R = e^r_cc - 1
    # r_cc = ln(R + 1)
    r_A = log(R_A + 1)
    r_C = log(R_C + 1)

    r_A
    r_C
    
```

13\. Suppose Amazon paid a $0.10 per share cash dividend at the end of October. What
is the monthly simple total return on Amazon? What is the monthly dividend yield?


```{r}
    
    # simple total return = (pt1 - pt0 + div) / pt0
    div_A = 0.10
    R_A_tot = (pA1 - pA0 + div_A)/ pA0
    R_A_tot
    
    # div yield = div / pt0 
    div_A / pA0


```

14\. Suppose the monthly returns on Amazon from question 12 above are the same
every month for 1 year. Compute the simple annual returns, as well as the
continuously compounded annual returns.


```{r}

    # (1 + R_annual) = (1 + R_monthly)^12
    R_A_12 = (1 + R_A)^12 - 1
    R_A_12

    # r_A_cc = ln(1 + R_A)
    log(1 + R_A_12)

```

15\. At the end of September 2004, supposed you have $10,000 to invest in
Amazon and Costco over the next month. If you invest $8,000 in Amazon and
$2,000 in Costco, what are your portfolio shares, xA and xC. 


```{r}

    xA = 8/10
    xC = 2/10

```

16\. Continuing with the previous question, compute the monthly simple return
on the portfolio. Assume than no dividends are paid.


```{r}

    # R_portfolio = w1 * R1 + w2 * R2 + ...
    R_p = xA * R_A + xC * R_C
    R_p

```

