
# Coursera: Intro to Computational Finance: Assignment 9-R: Portfolio Analysis

## Load Data

[http://s3.amazonaws.com/assets.datacamp.com/course/compfin/lab9.RData](http://s3.amazonaws.com/assets.datacamp.com/course/compfin/lab9.RData)

Copied locally (binary file): [../lab9.RData](lab9.RData)

```{r}

    library("zoo")

    # defines returns_df in the workspace
    load("lab9.RData")
    
    str(returns_df)

    rbind(head(returns_df),
          tail(returns_df))


    mu.vec <- apply(returns_df, 2, mean)
    sigma.vec <- apply(returns_df, 2, sd)
    Sigma.mat <- cov(returns_df)
    Rho.mat <- cor(returns_df)

    returns_df.mat <- coredata(returns_df)

    pairs(returns_df.mat,
          col="blue",
          pch=16)

    rbind(mu.vec, sigma.vec)

    rbind(Sigma.mat,Rho.mat)


```


### Question 1  Which two assets have the highest correlation?

```{r}

    Rho.mat

```


### Question 2  What is the weight of Microsoft in the global minimum variance portfolio?

```{r}

    source("../my.portfolio.r")
    
    gmv.portfolio.weights = gmvPortfolio(mu.vec, Sigma.mat)
    gmv.portfolio.weights

```

### Question 3  What is the standard deviation of the global minimum variance portfolio?

```{r}

    gmv.portfolio.mu = t(gmv.portfolio.weights) %*% mu.vec
    gmv.portfolio.sigma =  sqrt( t(gmv.portfolio.weights) %*% Sigma.mat %*% gmv.portfolio.weights )

    c("mu"=gmv.portfolio.mu, "sigma"=gmv.portfolio.sigma)


```

### Question 4  What is the expected return of the global minimum variance portfolio?


```{r}

    c("mu"=gmv.portfolio.mu, "sigma"=gmv.portfolio.sigma)

```

### Question 5  What happens to the global minimum variance portfolio if short sales are restricted?

```{r}

    gmv.ns.portfolio.weights <- gmvPortfolio.noShort(mu.vec, Sigma.mat)

    rbind( gmv.portfolio.weights, gmv.ns.portfolio.weights)

```

### Question 6  Compute Efficient Portfolio with Target Return Equal to the MAX Individual Asset Return:

Of the four stocks, determine the stock with the largest estimated expected return. 
Use this maximum average return as the target return for the computation of an
efficient portfolio allowing for short-sales. 

What is the weight of Microsoft in this portfolio?

```{r}

    eff.portfolio.weights <- effPortfolio(mu.vec, Sigma.mat, max(mu.vec))
    eff.portfolio.weights 

```

### Question 7  Compute Efficient Portfolio - NO SHORTING - with Target Return Equal to the MAX Individual Asset Return:


Of the four stocks, determine the stock with the largest estimated expected return. 
Use this maximum average return as the target return for the computation of an
efficient portfolio not allowing for short-sales. What is the weight of
Microsoft in this portfolio?


```{r}

    eff.ns.portfolio.weights <- effPortfolio.noShort(mu.vec, Sigma.mat, max(mu.vec))
    eff.ns.portfolio.weights 

```


### Question 8 Convex combinations of efficient portfolios:

Using the fact that all efficient portfolios can be written as a convex
combination of two efficient portfolios, compute efficient portfolios as convex
combinations of the global minimum variance portfolio and the efficient
portfolio that was computed in question six. 

What is the expected return of the portfolio when α=.5?

```{r}

    alpha.vec=seq(0,1,by=0.1)
    eff.frontier.weights <- effFrontier(mu.vec, 
                                        Sigma.mat,
                                        alpha.vec=alpha.vec,
                                        gmv.portfolio.weights=gmv.portfolio.weights,
                                        eff.portfolio.weights=eff.portfolio.weights)
    cbind(alpha.vec, eff.frontier.weights)

    eff.frontier.5.weights <- eff.frontier.weights[6,]
    eff.frontier.5.mu <- eff.frontier.5.weights %*% mu.vec
    eff.frontier.5.mu


```

### Question 9  What is the weight of Microsoft in the tangency portfolio with short sales allowed?


```{r}
    
    rf <- 0.005
    tan.portfolio.weights <- tanPortfolio(mu.vec, Sigma.mat, rf)
    tan.portfolio.weights


```

### Question 10  What is the weight of Microsoft in the tangency portfolio with short sales not allowed?

```{r}

    tan.ns.portfolio.weights <- tanPortfolio.noShort(mu.vec, Sigma.mat, rf)
    tan.ns.portfolio.weights

```
